### 2016-06-29 publications/Journal/1-introduction.tex
Revision, a few modifications in 1,5,6
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

This paper introduces DEVS-Ex-Machina\footnote{\url{https://bitbucket.org/bcardoen/devs-ex-machina}} (``dxex''): our simulation tool~\cite{dxex} which offers multiple synchronization protocols: no synchronization (sequential execution), conservative synchronization, or optimistic synchronization.
The selected synchronization protocol is transparent to the simulated model: users only determine which protocol to use.
Users who simulate a wide variety of models, with different ideal synchronization protocols, can run the same model with different synchronization protocols.
Since the ideal synchronization protocol might change throughout the simulation, we also support runtime switching between these protocols.
This runtime switching is based on performance metrics, which are logged during simulation.
Information is made available to a seperate component, where a choice is made about which synchronization protocol to use next.
Additionally, we investigate in this paper how model allocation influences the performance of our synchronization protocols.
To this end, we have included an allocation component to our simulation kernel.

@@

We implemented a model that, depending on a single parameter, changes its ideal synchronization protocol.
Using several models, we demonstrate the factors influencing the performance under a given synhronization protocol.
Dxex, then, is used to compare simulation using exactly the same tool, but with a varying synchronization protocol.
With dxex users can always opt to use the fastest protocol available, and through its modularity, users could even implement their own.
To verify that this modularity does not hinder performance, we compare to another Parallel DEVS simulation kernel, called adevs~\cite{adevs}.

>>>>

This paper introduces DEVS-Ex-Machina\footnote{\url{https://bitbucket.org/bcardoen/devs-ex-machina}} (``dxex''): our simulation tool~\cite{dxex} which offers multiple synchronization protocols: no synchronization (sequential execution), conservative synchronization, or optimistic synchronization.
The selected synchronization protocol is transparent to the simulated model: users only determine which protocol to use.
Users who simulate a wide variety of models, with different ideal synchronization protocols, can run the same model with different synchronization protocols.
Since the ideal synchronization protocol might change throughout the simulation, we also support runtime switching between these protocols.
This runtime switching can be based on performance metrics, which are logged during simulation.
Information is made available to a separate component, where a choice can be made about which synchronization protocol to use next. % It's manual now, can be extended to full automatic.
Additionally, we investigate in this paper how model allocation influences the performance of our synchronization protocols.
To this end, we have included an allocation component to our simulation kernel.

@@

We implemented a model that, depending on a single parameter, changes its ideal synchronization protocol.
Using several models, we demonstrate the factors influencing the performance under a given synchronization protocol.
Dxex, then, is used to compare simulation using exactly the same tool, but with a varying synchronization protocol.
With dxex users can always opt to use the fastest protocol available, and through its modularity, users could even implement their own.
To verify that this modularity does not hinder performance, we compare to another Parallel DEVS simulation kernel, called adevs~\cite{adevs}.

### 2016-06-29 publications/Journal/6-allocation.tex
Finished going through the citations
Revision, a few modifications in 1,5,6
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

The PHOLDTree model is similar in structure to models of gossiping in social networks~\cite{Gossip} and multi-granularity locking~\cite{MultiGranularityLocking}.
The lookahead of an atomic node is the minimally allowed $\epsilon$, indicating uncertainty, as is often the case in realistic models.
We demonstrate the importance of allocation by comparing performance of a breadth-first versus a depth-first scheme.
Both options are automated ways of allocation that are independent of the model.

@@

After simulation the traces can be visualized for both breadth-first and depth-first allocation.
Using a breadth-first allocation scheme, as shown in Figure~\ref{fig:pholdtree_visualize_parBFS}, we notice that many events get exchanged between cores.
This is caused by the high number of inter-core connections and the high number of events exchanged over these connections.
The number of connections between nodes at the same simulation core is also rather low.
Using a depth-first allocation scheme, however, as shown in Figure~\ref{fig:pholdtree_visualize_parDFS},

@@

The most prominent aspect of these results is the low performance for conservative depth-first allocation for two cores.
This is mostly caused by the difference between a sequential simulation and a parallel simulation: suddenly we need to take into account synchronization and passing around of lookahead values.
And since the number of cores is low, the overhead dominates any further speedup.
%TODO add some more explanation here maybe?
%TODO why is this not the case for optimistic?

%TODO this paragraph is strange: is this really the case?
Interestingly, we see that optimistic synchronization is less influenced by the allocation than conservative synchronization.
This is likely caused due to the lower number of connections to take into account in conservative synchronization.
Whereas conservative synchronization needs to take into account even scarcely used connections, optimistic synchronization does not.
The same is true in the opposite direction, though, where optimistic synchronization is slower when a good allocation is chosen.
Conservative synchronization will then be able to make better estimates, whereas optimistic synchronization does not make estimations.

>>>>

The PHOLDTree model is similar in structure to models of gossiping in social networks~\cite{Gossip}.
The lookahead of an atomic node is the minimally allowed $\epsilon$, indicating uncertainty, as is often the case in realistic models.
We demonstrate the importance of allocation by comparing performance of a breadth-first versus a depth-first scheme.
Both options are automated ways of allocation that are independent of the model.

@@

After simulation the traces can be visualized for both breadth-first and depth-first allocation.
Using a breadth-first allocation scheme, as shown in Figure~\ref{fig:pholdtree_visualize_parBFS}, we notice that many events get exchanged between cores.
This is caused by the high number of inter-core connections and the high number of events exchanged over these connections.
The number of connections between nodes at the same simulation core is also rather low.
Using a depth-first allocation scheme, however, as shown in Figure~\ref{fig:pholdtree_visualize_parDFS}, minimizes inter-core connections while maximizing intra-core connections.

@@

The most prominent aspect of these results is the low performance for conservative depth-first allocation for two cores.
This is mostly caused by the difference between a sequential simulation and a parallel simulation: suddenly we need to take into account synchronization and passing around of lookahead values.
And since the number of cores is low, the overhead dominates any further speedup.
Optimistic is less sensitive to the number of models per core as it does not need to poll each model for a lookahead, this explains the lower runtime penalty observed for optimistic.
%TODO add some more explanation here maybe?
%TODO why is this not the case for optimistic?

%TODO this paragraph is strange: is this really the case? 
%The effect is due to the topology of the kernels, in BFS the risk on reverts is higher. Optimistic is less influenced but will get a lower speedup in the good case.
Interestingly, we see that optimistic synchronization is less influenced by the allocation than conservative synchronization.
This is likely caused due to the lower number of connections to take into account in conservative synchronization.
Whereas conservative synchronization needs to take into account even scarcely used connections, optimistic synchronization does not.
The same is true in the opposite direction, though, where optimistic synchronization is slower when a good allocation is chosen.
Conservative synchronization will then be able to make better estimates, whereas optimistic synchronization does not make estimations.
